Word2Vec Analysis

### 2018-08-03 Main.Rmd
Progress on clustering comparison.

---
title: "Replication of Hills 2009"
author: "Isaac Scheinfeld"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: default
---
# Background

...

# Networks

@@

full_vocab <- make_vocab_dataframe(lang="English (American)",
                              lang_form = "WS",
                              lex_class = "nouns") %>%
  trim_all_unilemma() %>%
  trim_all_definition() %>%
  arrange(age) %>%            # keep first age for duplicates
  distinct(uni_lemma, .keep_all = TRUE) %>%
  select(uni_lemma, definition, age)
```

@@

```{r warning=FALSE}
features <- load_features()

hills_vocab <- full_vocab %>%
  filter(age <= 30) %>%
  filter(uni_lemma %in% features$Concept) # filter to words with features
```

To begin our analysis, we only consider the words for which feature data are available, an overlap of `r nrow(vocab)` in our data compared to Hills' 130.

## Word2Vec

## Build Networks

Using vocabulary obtained by 30 months and the same McRae features as Hills, we construct networks from perceptual and conceptual features. We track the number of features that generate each edge.

```{r message=FALSE}
# perceptual and conceptual McRae features
perceptual_features = c("sound", "taste", "smell", "tactile",
                        "visual-form_and_surface", "visual-motion", "visual-color")
conceptual_features = c("function")

# count shared features between words in vocabulary
perceptual_links <- make_feature_links(vocab = hills_vocab,
                                       feature_types = perceptual_features) %>%
  rename(perceptual = shared)

conceptual_links <- make_feature_links(vocab = hills_vocab,
                                       feature_types = conceptual_features) %>%
  rename(conceptual = shared)

# get word2vec distance between words in vocabulary
w2v_links <- make_w2v_links(vocab = hills_vocab) #TODO still problem w/ vocab matching w2v data

@@

full_network <- graph_from_data_frame(links, vertices = hills_vocab, directed = FALSE) %>%
  as_tbl_graph()
```

@@

## The Full Network

### Network statistics and small-world analyses

```{r}
# testing dense layout
source(paste(getwd(),"/helpers/all_helper.r",sep = ""), chdir = T)

plot_network(full_network %E>% filter(w2v_similarity >= 0.5), remove_isolated = TRUE, labels=TRUE, frame=TRUE)
```

The full network is shown in the following figure, for w = 1, 2, 3, and 4.

```{r out.width = '100%'}
full_networks = list()
for (w in 1:4) {
  full_networks[[w]] = full_network %E>%
    filter(shared >= w)
}

#par(mfcol=c(2, 2), mar=c(0.5,0.5,0.5,0.5))
for (network in full_networks) {
  plot_network(network, remove_isolated = FALSE, labels=TRUE, frame=TRUE)
}
```

## Basic Vocab and Feature Properties

<!-- ## Basic Network Properties -->

<!-- Some basic properties of networks built from perceptual and conceptual features. -->
<!-- ```{r, results='asis'} -->
<!-- max_w <- 4 -->

<!-- network_properties <- function(networks) { -->
<!--   return( tibble( -->
<!--   name = map_chr(networks, ~ .x$name), -->
<!--   type = map_chr(networks, ~ .x$type), -->
<!--   age = map_dbl(networks, ~ .x$age) %>% -->
<!--     as.integer(), -->
<!--   w = map_dbl(networks, ~ .x$w) %>% -->
<!--     as.integer(), -->
<!--   vertices = map_int(networks, vcount), -->
<!--   edges = map_dbl(networks, ecount) %>% -->
<!--     as.integer(), # for some reason igraph ecount returns double -->
<!--   avg_degree = map(networks, degree) %>% -->
<!--     map_dbl(mean), -->
<!--   avg_shortest_path = map_dbl(networks, mean_distance), -->
<!--   diameter = map_dbl(networks, diameter) %>% -->
<!--     as.integer(), -->
<!--   clustering_coefficient = map_dbl(networks, transitivity), -->
<!--   density = edges / (vertices * (vertices - 1) / 2), -->
<!--   avg_shortest_path_random = 0, # either estimate or calculate values for random nets -->
<!--   diameter_random = 0, -->
<!--   clustering_coefficient_random = density, -->
<!--   isolates = map_dbl(networks, function (g) sum(degree(g) == 0)), -->
<!--   components = map_dbl(networks, count_components) %>% -->
<!--     as.integer() - isolates)) -->
<!-- } -->

<!-- full_networks <- list() -->
<!-- perceptual_networks <- list() -->
<!-- conceptual_networks <- list() -->
<!-- for (w in 1:max_w) { -->
<!--   full_networks[[w]] <- full_network %>% -->
<!--     activate(what = edges) %>% -->
<!--     filter(shared >= w) %>% -->
<!--     set.graph.attribute("w", w) -->

<!--   perceptual_networks[[w]] <- perceptual_network %>% -->
<!--     activate(what = edges) %>% -->
<!--     filter(shared >= w) %>% -->
<!--     set.graph.attribute("w", w) -->

<!--   conceptual_networks[[w]] <- conceptual_network %>% -->
<!--     activate(what = edges) %>% -->
<!--     filter(shared >= w) %>% -->
<!--     set.graph.attribute("w", w) -->
<!-- } -->

<!-- networks <- full_networks %>% -->
<!--   append(perceptual_networks) %>% -->
<!--   append(conceptual_networks) -->

<!-- our_network_properties <- network_properties(networks) -->

<!-- our_network_properties %>% -->
<!--   select(type, w, vertices, edges, avg_degree, diameter, avg_shortest_path, clustering_coefficient) %>% -->
<!--   kable(format="html", -->
<!--         escape = FALSE, -->
<!--         col.names = c("Type", "w", "V", "E", "Average degree", "Diameter", "Average length", "Clustering coefficient"), -->
<!--         digits = 2, -->
<!--         align = rep("l", 8)) %>% -->
<!--   kable_styling(full_width = F) %>% -->
<!--   collapse_rows(columns = 1:3, valign = "top") -->
<!-- ``` -->

<!-- We can compare these networks to those from Hills, noting that we have a different set of words/nodes but that the induced subgraph of those words shared between our analysis and that of Hills should be the same. -->
<!-- ```{r} -->
<!-- hills_network_properties <- tibble( # table from Hills 2009 -->
<!--   type = c(rep("all", times=4), rep("perceptual", times=4), rep("conceptual", 4)), -->
<!--   w = c(1:max_w, 1:max_w, 1:max_w), -->
<!--   clustering_coefficient = c(0.55, 0.54, 0.58, 0.60, 0.54, 0.54, 0.62, 0.62, 0.88, 0.59, 0.38, 1.00), -->
<!--   avg_shortest_path = c(1.72, 2.57, 4.96, 3.22, 1.76, 2.69, 5.02, 2.55, 3.08, 1.59, 5.02, 1.00), -->
<!--   density = c(0.29, 0.10, 0.04, 0.02, 0.27, 0.08, 0.03, 0.01, 0.05, 0.01, 0.0, 0.0), -->
<!--   components = c(1, 1, 4, 10, 1, 2, 8, 11, 7, 14, 7, 2), -->
<!--   isolates = c(0, 6, 20, 47, 0, 10, 31, 69, 33, 81, 113, 126) -->
<!-- ) -->

<!-- # convert contents to text -->
<!-- display_our_properties <- our_network_properties %>% mutate_all(funs(format(., digits=2, trim=TRUE))) -->
<!-- display_hills_properties <- hills_network_properties %>% mutate_all(funs(paste("(", ., ")", sep=''))) -->

<!-- # combine text tables for display -->
<!-- display_properties <- bind_rows(display_our_properties %>% mutate(row = 1:n()), -->
<!--                                 display_hills_properties %>% mutate(row = 1:n())) %>% -->
<!--   group_by(row) %>% -->
<!--   summarise_if(is.character, funs(paste(., collapse=" "))) %>% -->
<!--   arrange(row) -->

<!-- # display combined properties -->
<!-- select(hills_network_properties, type, w) %>% -->
<!--   bind_cols(select(display_properties, clustering_coefficient, avg_shortest_path, -->
<!--                    density, components, isolates)) %>% -->
<!--   kable(format="html", -->
<!--         escape = FALSE, -->
<!--         col.names = c("Type", "w", "Clustering Coefficient", "Average length", "Density", "Components", "Isolates"), -->
<!--         align = rep("l", 7)) %>% -->
<!--   kable_styling(full_width = F) %>% -->
<!--   collapse_rows(columns = 1:2, valign = "top") -->
<!-- ``` -->

<!-- # Clusters -->

<!-- ## Visualize Networks -->
<!-- ```{r out.width = '100%'} -->
<!-- networks <- perceptual_networks[2] -->
<!-- ``` -->

<!-- ## Clique Percolation -->

<!-- Initial Clique Analysis -->
<!-- ```{r out.width = '100%'} -->
<!-- trimmed_networks <- map(networks, function (g) delete.vertices(g, degree(g)==0)) -->

<!-- # cl <- makeCluster(3) -->
<!-- # registerDoParallel(cl) -->
<!-- # getDoParWorkers() -->
<!-- ## Plot with the new colors -->
<!-- for (net in networks) { -->
<!--   for (k in 3:4) { -->
<!--     clusters <- clique.community.opt(net,k) -->
<!--     plot_network(net, clusters = clusters, labels=FALSE) -->
<!--   } -->
<!-- } -->
<!-- # stopCluster(cl) -->
<!-- ``` -->

<!-- Dynamic Visualization -->
<!-- ```{r} -->
<!-- conceptual_clusters <- membership(cluster_walktrap(conceptual_network)) -->

<!-- # Convert to object suitable for networkD3 -->
<!-- conceptual_d3 <- igraph_to_networkD3(conceptual_network, group = conceptual_clusters) -->

<!-- # Create force directed network plot -->
<!-- forceNetwork(Links = conceptual_d3$links, Nodes = conceptual_d3$nodes, -->
<!--              Source = 'source', Target = 'target', NodeID = 'name', -->
<!--              Group = 'group') -->
<!-- ``` -->

>>>>

---
title: "Word2Vec Analysis"
author: "Isaac Scheinfeld"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: default
---
# Networks

@@

full_vocab <- make_vocab_dataframe(lang="English (American)",
                                   lang_form = "WS",
                                   lex_class = "nouns") %>%
  trim_all_unilemma() %>%
  trim_all_definition() %>%
  arrange(age) %>%            # keep first age for duplicates
  distinct(uni_lemma, .keep_all = TRUE) %>%
  select(uni_lemma, definition, age, category)
```

@@

```{r warning=FALSE, message=FALSE}
features <- load_features()
w2v <- load_w2v()

hills_vocab <- full_vocab %>%
  filter(age <= 30) %>%
  filter(uni_lemma %in% features$Concept) # filter to words with features

vocab <- hills_vocab %>%
  filter(uni_lemma %in% w2v$from | uni_lemma %in% w2v$to)
```

To begin our analysis, we only consider the words for which feature data are available, an overlap of `r nrow(hills_vocab)` in our data compared to Hills' 130. Then we filter the `r nrow(hills_vocab) - nrow(vocab)` words for which word2vec data is not available, for a final vocabulary of `r nrow(vocab)`.

## Build Network

Using vocabulary obtained by 30 months and the same McRae features as Hills, we construct a network with edge attributes for perceptual and conceptual features, and for word2vec similarity.

```{r message=FALSE}
# perceptual and conceptual McRae feature types
perceptual_features = c("sound", "taste", "smell", "tactile",
                        "visual-form_and_surface", "visual-motion", "visual-color")
conceptual_features = c("function")

# count shared features between words in vocabulary
perceptual_links <- make_feature_links(vocab = vocab,
                                       feature_types = perceptual_features) %>%
  rename(perceptual = shared)

conceptual_links <- make_feature_links(vocab = vocab,
                                       feature_types = conceptual_features) %>%
  rename(conceptual = shared)

# get word2vec distance between words in vocabulary
w2v_links <- make_w2v_links(vocab = vocab) #TODO still problem w/ vocab matching w2v data

@@

network <- graph_from_data_frame(links, vertices = vocab, directed = FALSE) %>%
  as_tbl_graph()
```

@@

## The Network

First we generate a display layout for the network, based on the word2vec data.

```{r}
source(paste(getwd(),"/helpers/all_helper.r",sep = ""), chdir = T)

layout <- layout_network(network %E>% mutate(weights = w2v_similarity))
```

We can examine the networks generated by filtering edges with a word2vec similarity threshold.

```{r out.width = '100%', fig.height = 5, fig.width = 5}
source(paste(getwd(),"/helpers/all_helper.r",sep = ""), chdir = T)

par(mfrow=c(2, 2), mar=c(2,0.4,0.8,0.4)) # row x col, column first

for (s in seq(0.35, 0.5, 0.05)) {
  plot_network(network %E>% filter(w2v_similarity > s),
               layout = layout,
               labels=TRUE,
               title=str_glue("Threshold = {s}"))
}

```

Test igraph clustering algorithms

```{r fig.height = 5, fig.width = 5}
source(paste(getwd(),"/helpers/all_helper.r",sep = ""), chdir = T)

# cluster_edge_betweenness - doesn't cluster reasonably, treats weights as distances?

# cluster_infomap - returns one cluster

# cluster_label_prop - returns one cluster

# cluster_optimal - too computationally intensive for laptop, may not be possible (NP)

plot_network(network, layout = layout,
             clusters = cluster_fast_greedy(network, weights = E(network)$w2v_similarity)$membership,
             labels=TRUE, frame=TRUE)

plot_network(network, layout = layout,
             clusters = cluster_leading_eigen(network, weights = E(network)$w2v_similarity)$membership,
             labels=TRUE, frame=TRUE)

plot_network(network, layout = layout,
             clusters = cluster_louvain(network, weights = E(network)$w2v_similarity)$membership,
             labels=TRUE, frame=TRUE)

plot_network(network, layout = layout,
             clusters = cluster_spinglass(network, weights = E(network)$w2v_similarity)$membership,
             labels=TRUE, frame=TRUE)

plot_network(network, layout = layout,
             clusters = cluster_walktrap(network, weights = E(network)$w2v_similarity)$membership,
             labels=TRUE, frame=TRUE)
```
